---
title: "XAI3_v2"
author: "Hajar Abdoun"
date: "13/5/2022"
output: html_document
---

## EXERCISE:

Apply PDP to the regression example of predicting bike rentals. Fit a random forest approximation for the prediction of bike rentals (**cnt**). Use the partial dependence plot to visualize the relationships the model learned. Use the slides shown in class as model.  

## QUESTION:

Analyse the influence of **days since 2011, temperature, humidity** and **wind speed** on the predicted bike counts.


```{r}
library(dplyr)
library(plotly)
library(reshape2)
library(lubridate)
library(randomForestSRC)

#setwd("/Users/cmonserr/OneDrive - UPV/Trabajo_2/Asignaturas/Evaluacion de modelos/Practicas/Practica 3/Bike-Sharing-Dataset")
days <- read.csv("day.csv")
hour <- read.csv("hour.csv")

days$dteday <- as_date(days$dteday)
days_since <- select(days, workingday, holiday, temp, hum, windspeed, cnt)
days_since$days_since_2011 <- int_length(interval(ymd("2011-01-01"), days$dteday)) / (3600*24)
days_since$SUMMER <- ifelse(days$season == 3, 1, 0)
days_since$FALL <- ifelse(days$season == 4, 1, 0)
days_since$WINTER <- ifelse(days$season == 1, 1, 0)
days_since$MISTY <- ifelse(days$weathersit == 2, 1, 0)
days_since$RAIN <- ifelse(days$weathersit == 3 | days$weathersit == 4, 1, 0)
days_since$temp <- days_since$temp * 47 - 8
days_since$hum <- days_since$hum * 100
days_since$windspeed <- days_since$windspeed * 67

rf <- rfsrc(cnt~., data=days_since)

results <- select(days_since, days_since_2011, temp, hum, windspeed, cnt)
nr <- nrow(days_since)
for(c in names(results)[1:4])
{
  for(i in 1:nr){
    r <- days_since
    r[[c]] <- days_since[[c]][i]
    sal <- predict(rf, r)$predicted
    results[[c]][i] <- sum(sal) / nr
  }
}

p1 = ggplot(days_since, aes(x = temp, y = results$temp)) + ylim(0, 6000) + geom_line() + geom_rug(sides="b", alpha = 0.5) + labs(x = "Temperature")

p2 = ggplot(days_since, aes(x = hum, y = results$hum)) + ylim(0, 6000) + geom_line() + geom_rug(sides="b", alpha = 0.5) + labs(x = "Humidity")

p3 = ggplot(days_since, aes(x = days_since_2011, y = results$days_since_2011)) + ylim(0, 6000) + geom_line() + geom_rug(sides="b", alpha = 0.5) + labs(x = "Days since 2011", y = "Predictions")

p4 = ggplot(days_since, aes(x = windspeed, y = results$windspeed)) + ylim(0, 6000) + geom_line() + geom_rug(sides="b", alpha = 0.5) + labs(x = "Windspeed")
#p1
subplot(p3, p1, p2, p4, titleX = TRUE, titleY = TRUE, shareX = FALSE, shareY = TRUE)
```
Days since 2011: as time goes by, there is generally a clear upward trend in bicycle rentals, although in the last few days the predictions are decreasing. 

Temperature: The number of rented bicycles increases with the temperature up to around 20 degrees Celsius, but then decreases as the temperature rises from around 25 degrees Celsius onwards. 

Humidity: up to 50% humidity, the number of rented bicycles remains constant, but then decreases inversely to the increase in humidity. 

Windspeed: the prediction for bicycle rentals decreases gradually until a wind speed of approximately 23 km/h is reached. At this point, the predictions are constant.


## EXERCISE:

Generate a 2D Partial Dependency Plot with humidity and temperature to predict the number of bikes rented depending of those parameters.

BE CAREFUL: due to the size, extract a set of random samples from the BBDD before generating the the data for the Partial Dependency Plot. 

Show the density distribution of both input features with the 2D plot as shown in the class slides. 

TIP: Use geom_tile() to generate the 2D plot. Set width and height to avoid holes. 

## QUESTION:

Interpret the results.


```{r}

sampled <- sample_n(days_since, 40)
temp <- sampled$temp
hum <- sampled$hum
th <- inner_join(data.frame(temp),data.frame(hum), by=character())
th$p <- 0

for(i in 1:nrow(th)){
  r <- days_since
  r[["temp"]] <- th[["temp"]][i]
  r[["hum"]] <- th[["hum"]][i]
  
  sal <- predict(rf, r)$predicted
  th[["p"]][i] <- sum(sal) / nr
}

#View(th)
```

```{r}
p5 = ggplot(th, aes(x = temp, y = hum, fill = p)) + geom_tile(width = 10, height = 15) + geom_rug(alpha = 0.5) + labs(x = "Temperature", y = "Humidity", fill = "Number of bikes")

p5
```
In this graph we can see that the highest number of bicycles are rented when the temperature is between 15 and 20 degrees and the relative humidity is between 0 and 70%, while the lowest number of bicycles are rented when the humidity is very high and the temperature is very low. 